---
title: "Cell Count analysis 11-16"
output: html_notebook
---

The data taken on 11-16 which was the week before thanksgiving.  Prior to these cells being exposed to the magnetic field, the incubator ran out of CO2.  This happened Wednesday morning which was the day of the trials.  I ended up moving the cells to Ana's incubator in the morning and then taking them out again later that afternoon for the magnetic field exposure. 

These cells were seeded Monday afternoon, left to grow on Tuesday, and then experimental exposure started on Wednesday. Counts were performed Wednesday before the exposure and during exposure, cells that were not undergoing the treatment were placed in the incubator.

```{r}
# clear the environment window
rm(list = ls())

# include the libraries that we need
library(ggplot2)
library(dplyr)
library(here)
library(lubridate)
library(ggfortify)
```

Now we will import the dataset

```{r}
cells <- read.csv(here("Data", "Cellcountdata_11-16_11-18.csv"), stringsAsFactors = TRUE)
```


Now let's take a look at the spread of the data with a boxplot real quick
```{r}
# relevel the data to put the control first
cells<- mutate(cells, Condition = relevel(Condition, ref = "Control"))

levels(cells$Condition)
```

```{r}
# reassign the values of the different EMF conditions to combine them
cells <- cells %>% mutate(Exposure = recode(cells$Date, "11/17/2022" = 'After Exposure', "11/16/2022" = 'Before Exposure'))

# relevel the data to put the before exposure first
cells<- mutate(cells, Exposure = relevel(Exposure, ref = "Before Exposure"))


# make a box plot of the data
ggplot(cells, aes(x = Condition, y = Cells)) +
geom_boxplot()+
  geom_point(size = 3, colour = 'lightgrey', alpha = 0.5)+
  facet_wrap(~Exposure)+
  xlab("Sample") +
  ylab("Cell counts (cells/mm^2)")
```
From this, it looks like the CO2 deprivation did impact the cells quite a bit and it doesn't look like they bounced back as quick as I thought they would

Let's make a scatterplot to visualize this a little better.
them on a new graph that is a little prettier. 


```{r}
# find the means of the cell counts and put them in a new table
# find the means of the cell counts and put them in a new table
mean_counts <- cells %>% group_by(Condition,Exposure) %>% summarise(
  meancounts = mean(Cells),
  secounts = sd(Cells)/sqrt(sum(!is.na(Cells))                                                             ))
```

Now let's plot this:
```{r}
# create scatterplot with error bars
ggplot(mean_counts, aes(Condition, meancounts, color = Exposure))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    facet_wrap(~Exposure)+
    theme(legend.position="none")+
    xlab("Condition")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
```

Since a statistically significant number of counts were performed for this condition, we will do a two sample t test to figure out the differences between them.

```{r}
# filter based on exposure
before <- cells %>% filter(Exposure == "Before Exposure")

after <- cells %>% filter(Exposure == "After Exposure")
  
  
# perform two sample t test
t.test(Cells ~ Condition, data = before)

t.test(Cells ~ Condition, data = after)
```

