---
title: "Mouse Hippocampal Cell Count Data Analysis for 11-1"
output: html_notebook
---

This code is designed to analyze the cell count data from the first experimental trial after putting mouse hippocampal cells in the EMF for 30 mins.  The goal of this trial is to measure survivalbility of these cells and ensure that the current protocol won't kill these neural cells before we try with C17.2 stem cells. 

```{r}
# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)

```


```{r}
# import the data set that we will be analysing, the stringsAsFactors = True argument converts categorical data to factors so we can analyse the levels
counts <- read.csv(here("Data", "MH_Cellcountdata_11-1.csv"),stringsAsFactors = TRUE)

# take a look at the structure of the data
glimpse(counts)
```

```{r}
# convert the count trial to factor data
counts$Count_trial<-as.factor(counts$Count_trial) 


# select the first two counts (before and after the EMF exposure) because the third count is skewed due to the lack of CO2 that the cells were exposed to between the 2nd and 3rd counts

counts <- counts %>% filter(Count_trial == c("1", "2"))
```

Now I am going to create a box plot of the data to visualize the spread and get a better idea of the distribution.
```{r}
# make a box plot of the data
ggplot(counts, aes(x = Condition, y = Counts)) +
geom_boxplot()+
  geom_point(size = 3, colour = 'lightgrey', alpha = 0.5)+
  facet_wrap(~Count_trial)+
  xlab("Sample") +
  ylab("Cell counts (cells/mm^2)")
```




