---
title: "Mouse Hippocampal Cell Count Data Analysis"
output: html_notebook
---

This code is designed to analyze the cell count data from the first experimental trial after putting mouse hippocampal cells in the EMF for 30 mins.  The goal of this trial is to measure survivalbility of these cells and ensure that the current protocol won't kill these neural cells before we try with C17.2 stem cells. 

```{r}
# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)

```


```{r}
# import the data set that we will be analysing, the stringsAsFactors = True argument converts categorical data to factors so we can analyse the levels
counts <- read.csv(here("Data", "MH_Cellcountdata_11-1.csv"),stringsAsFactors = TRUE)

# take a look at the structure of the data
glimpse(counts)
```

```{r}
# convert the sample to factor data

counts$Sample<-as.factor(counts$Sample) #convert to factor data type
levels(counts$Sample) #view levels of factor
```

Now I am going to create a boxplot of the data to visualize the spread and get a better idea of the distribution.
```{r}
# make a box plot of the data
ggplot(counts, aes(x = Sample, y = Counts)) +
geom_boxplot()+
  geom_point(size = 3, colour = 'lightgrey', alpha = 0.5)+
  facet_wrap(~Day)
  ylab("Sample") +
  xlab("Cell count (cells/mm^2")
```




