---
title: "Peltier and Thermistor Temperature testing 10-30"
output: html_notebook
---

For this trial I made some modifications to the Arduino control system. Instead of increasing based on the starting temperature of the peltier.  It is set to increase at a constant rate for 8 seconds before the thermistor turns on.  When the thermistor turns on, it starts modulating the temperature (for which I bumped down the threshold).  Let's see how this impacted the data between the control condition and the magnetic field condition.

```{r}
# clear R's environment window
rm(list = ls())

# import the libraries that we will need
library(ggplot2)
library(dplyr)
library(here)
```

```{r}
# import the data set
temp <- read.csv(here("Data", "Peltiertempanalysis10-30.csv"))

# open the dataset to check and make sure it is the correct one
View(temp)
```

```{r}
# plot a scatter plot that will give us an idea of the differences between the two conditions

ggplot(temp,
       aes(x = Time_sec, y = Avg, colour = Type))+
     geom_point()+
     xlab("Time (sec)")+
     ylab("Temperature (C)")+
     theme_bw()
```

