---
title: "CMF4 HT22 Cell Count Analysis 3-3 to 3-10"
author: "Cora Ferguson"
date: "2023-03-05"
output: html_document
---

This code is designed to analyze the cell count data from the second experimental trial after putting mouse hippocampal cells in the EMF for 30 mins.  The goal of this trial is to measure survivalbility of these cells and optimize the magnetic field exposure protocol.

Things that are different about this trial:
- cells seeded at 10,000 cells/plate
- 2 plates running at once (cells are swapped halfway through exposure)


```{r}
# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)

```

Import the data:
```{r}
# import the data set that we will be analysing, the stringsAsFactors = True argument converts categorical data to factors so we can analyse the levels
counts <- read.csv(here("Data", "CMF4_HT22Cells_3-3_3-10.csv"),stringsAsFactors = TRUE)

# take a look at the structure of the data
glimpse(counts)

# recode the counts to be before and after exposure
counts$Condition <- recode(counts$Condition, "B-Field" = 'Magnetic Field')

# relevel the data to put the control first
counts<- mutate(counts, Condition = relevel(Condition, ref = "Control"),
                        )



```

Now I am going to make a boxplot of the data to visualize the distributions
better idea of the distribution.

```{r}
# make a box plot of the data
ggplot(counts, aes(x = Condition, y = Cell_Count)) +
geom_boxplot()+
  geom_point(size = 3, colour = 'lightgrey', alpha = 0.5)+
  facet_wrap(~Trial_Day)+
  xlab("Sample") +
  ylab("Cell counts (cells/mm^2)")
```

The boxplot doesn't look the best but right now it is taking all of the count data and showing the distribution.  A more accurate teller of the actual data will be if we take the average of the counts by day.

```{r}
# find the means of the cell counts and put them in a new table
mean_counts <- counts %>% 
  filter(!is.na(Date), !is.na(Trial_Day), Condition == "Control" | Condition == "Magnetic Field") %>%
    group_by(Condition,Trial_Day) %>% summarise(
      meancounts = mean(Cell_Count),
      secounts = sd(Cell_Count)/sqrt(sum(!is.na(Cell_Count)))
  
       )
```

Great, now lets get to work with this data

```{r}
# create scatterplot with error bars
ggplot(mean_counts, aes(Condition, meancounts, color = Trial_Day))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    facet_wrap(~Trial_Day)+
    theme(legend.position="none")+
    xlab("Condition")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
         
```

Now lets make the relationship a little more obvious between the two

```{r}
library(scales)
# create scatter plot with error bars
ggplot(mean_counts, aes(Trial_Day, meancounts, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4', "dodgerblue"))+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    #theme(legend.position="none")+
    xlab("Trial Day")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
```
```{r}
# calculate the growth rates using a for loop and store the results of the data in   a new dataframe

#define the starting value of x (where we want the loop to start iterating through   our previous data set)
  x <- 1

# make an empty data frame
d <- data.frame()

# Write for loop to go through and create a new data set with the growth rates
  for(i in mean_counts$meancounts) {
    y = mean_counts$meancounts[x + 1]
    
    z = (y-mean_counts$meancounts[x])/24
    
    print(z)
    
    x = x+1
    
    d = rbind(d, z)
  }

```

```{r}
# add in the hours and slope variables to look at how the growth rate changes over time
mean_counts["Hours"] <-c(0,24,48,72,96, 120, 144, 0,24,48,72,96, 120, 144)
mean_counts["Slope"] <- c(NA,0.733, 0.866, NA, NA, NA, NA, NA,1.95, 1.478, NA, NA, NA, NA)

```

```{r}
# create scatter plot that shows the slopes of each of the conditions
ggplot(mean_counts, aes(Hours, Slope, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4'))+
     #geom_errorbar(aes(ymin = standardcounts - secounts,
                       #ymax = standardcounts + secounts), width = 0.1)+
    theme_bw()+
    #theme(legend.position="none")+
    xlab("Time (Hours)")+
    ylab(expression("Growthrate (Cells/hour)"))
```

Another way to visualize the growth rates:
```{r}
# create scatter plot that shows the slopes of each of the conditions
ggplot(mean_counts, aes(Hours, Slope, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4'))+
     #geom_errorbar(aes(ymin = standardcounts - secounts,
                       #ymax = standardcounts + secounts), width = 0.1)+
    geom_area(aes(fill = Condition, alpha = 0.2))+
    
    theme_bw()+
    scale_x_continuous(breaks=seq(0,144,by=24))+
    theme(legend.position="none")+
    xlab("Time (Hours)")+
    ylab(expression("Growthrate (Cells/hour)")) +
  facet_wrap(~Condition)
```

