---
title: "R Notebook"
output: html_notebook
---

Import the libraries that we need:
```{r}
# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)
library(readr)

```

Import the data:
```{r}
LDH1 <- read_csv(here("Data","CMF8_LDH_4-25.csv"))

LDH2 <- read_csv(here("Data", "CMF8_LDH_4-26.csv"))

LDH3 <- read_csv(here("Data",  "CMF8_LDH_4-27.csv"))
```
The first thing that I need to do is subtract the 680nm (noise) from the 490 absorbance reading.  I will combine all of the data frames together and then do that in one step

```{r}
# chop down LDH1 so we can bind these together
LDH1 <- LDH1 %>%
  select(1:7)

# combine the data frames together
full_LDH <- bind_rows("Day 1" = LDH1,
          "Day 2" = LDH2,
          "Day 3" = LDH3,
          .id = "Day")

```

Now lets subtract the 680 from the 490.  I will be doing this for the blanked and the raw data

```{r}
full_LDH_tidy <-
  full_LDH %>%
  mutate(Adjusted_Raw = Read_680 - Read_490,
         Adjusted_Blank = Blank_Read_680 - Blank_Read_490,
         Name = case_when(Name %in% c("Meida") ~ "Media"
        ,TRUE ~ Name),
         Name = case_when(Name %in% c("Control 1", "Control 2") ~ "Control Cells", TRUE ~ Name),
        Name = case_when(Name %in% c("BF 1", "BF 2")~"Magnetic Field Cells", TRUE ~ Name),
        Name = case_when(Name %in% c("BF Media") ~ "Magnetic Field Media", TRUE ~ Name),
        Name = case_when(Name %in% c("Positive")~ "Positive LDH Control", TRUE~Name)
        ) 
  #slice(-52, -98)
```

Alright, lets subset this to isolate the before and after absorbances of the raw and blanked data for each day:

```{r}
LDH_full_before <-
  full_LDH_tidy %>%
  filter(Well_ID == "BLK" | Well_ID == "CTL2" | Well_ID == "SPL1"| Well_ID == "SPL2" |Well_ID == "SPL3" | Well_ID == "SPL4" | Well_ID == "SPL5")

LDH_full_after <-
  full_LDH_tidy %>%
  filter(Well_ID == "BLK" | Well_ID == "CTL2" | Well_ID == "SPL6"| Well_ID == "SPL7" | Well_ID == "SPL8" | Well_ID == "SPL9" | Well_ID == "SPL 10")
```

Now lets try and graph these
```{r}
# calculate the means and the standard deviation of the data
LDH_before_means <-
  LDH_full_before %>%
  group_by(Day, Name) %>%
  summarise(mean_absorbance_raw = mean(Adjusted_Raw),
            se_absorbance_raw = sd(Adjusted_Raw)/sqrt(sum(!is.na(Adjusted_Raw))),
            mean_absorbance_blank = mean(Adjusted_Blank),
            se_absorbance_blank = sd(Adjusted_Blank)/sqrt(sum(!is.na(Adjusted_Blank)))) %>%
  mutate(mean_absorbance_raw = abs(mean_absorbance_raw),
         se_absorbance_raw = abs(se_absorbance_raw),
         mean_absorbance_blank = abs(mean_absorbance_blank),
         se_absorbance_blank = abs(se_absorbance_blank))



```

Graph for before exposure:
```{r}
ggplot(LDH_before_means, aes(Day, mean_absorbance_blank, color = Name, group = Name))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = mean_absorbance_blank - se_absorbance_blank,
                       ymax = mean_absorbance_blank + se_absorbance_blank), width = 0.1)+
    theme_bw()+
    labs(x = "Trial Day", y="Absorbance", color = "Sample")
```
Now lets repeat the process for after magnetic field exposure:
```{r}
# calculate the means and the standard deviation of the data
LDH_after_means <-
  LDH_full_after %>%
  group_by(Day, Name) %>%
  summarise(mean_absorbance_raw = mean(Adjusted_Raw),
            se_absorbance_raw = sd(Adjusted_Raw)/sqrt(sum(!is.na(Adjusted_Raw))),
            mean_absorbance_blank = mean(Adjusted_Blank),
            se_absorbance_blank = sd(Adjusted_Blank)/sqrt(sum(!is.na(Adjusted_Blank)))) %>%
  mutate(mean_absorbance_raw = abs(mean_absorbance_raw),
         se_absorbance_raw = abs(se_absorbance_raw),
         mean_absorbance_blank = abs(mean_absorbance_blank),
         se_absorbance_blank = abs(se_absorbance_blank))

```

Graph:
```{r}
ggplot(LDH_after_means, aes(Day, mean_absorbance_blank, color = Name, group = Name))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = mean_absorbance_blank - se_absorbance_blank,
                       ymax = mean_absorbance_blank + se_absorbance_blank), width = 0.1)+
    theme_bw()+
    labs(x = "Trial Day", y="Absorbance", color = "Sample")
```
Now lets take a look at the before and after absorbances for these (and seperate them by day):
```{r}
library(forcats)
# join the summary DFs
LDH_means <-
  bind_rows("Before" = LDH_before_means,
            "After" = LDH_after_means,
            .id = "Sample_Time") %>%
  mutate(Sample_Time = fct_relevel(Sample_Time, c("Before","After")))
```

Graph:
```{r}
ggplot(LDH_means, aes(Sample_Time, mean_absorbance_blank, color = Name, group = Name))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = mean_absorbance_blank - se_absorbance_blank,
                       ymax = mean_absorbance_blank + se_absorbance_blank), width = 0.1)+
  facet_wrap(~Day)+
    theme_bw()+
    labs(x = "Sample Time with Respect to Exposure", y="Absorbance", color = "Sample")
```




