---
title: "Microtubule aggregation analysis"
output: html_notebook
---

Another metric that we are interested in analyzing is the microtubule aggregation. The microtubules are a part of the cytoskeleton responsible for a lot of different things including structural integrity and organizing cell division.  Here, we are interested in seeing if there is a difference in the number of aggregations of microtubules between the conditions and the overall size of these aggregations.

```{r}
# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)
library(readr)
```

Import the data:
```{r}
control <- read_csv(here("Data", "controlMTaggregationresults.csv"))
  
bfield <- read_csv(here("Data", "bfieldMTaggregationresults.csv"))
```
Now let's bind these together and clean them up:
```{r}
full_df <- bind_rows("Control" = control,
          "Magnetic Field" = bfield,
          .id = "Condition")
```

Clean it up and isolate the length data we are interested in
```{r}
full_df_tidy <- full_df %>%
  select(Condition, Length) %>%
  mutate(Length = Length/2)
```

Now lets compare them:
```{r}
MT_test <- t.test(Length~Condition, data = full_df_tidy)
MT_test
```

Calculate the means and standard error:
```{r}
mean_MT <- full_df_tidy %>% 
    group_by(Condition) %>% summarise(
      meanlength = mean(Length),
      selength = sd(Length)/sqrt(sum(!is.na(Length))),
      n_observations = n()
  
       )
mean_MT
```

Now let's graph this:
```{r}
# create barplot with error bars
ggplot(full_df_tidy, aes(x = Condition, y = Length, fill = Condition))+
     geom_boxplot(alpha = 0.5)+
    theme_bw()+
    theme(legend.position="none")+
    xlab("Condition")+
    ylab("Width of Microtubule Aggregates(\u00b5m)")
     
```







