---
title: "CMF HT22 Microscopy Analysis"
output: html_notebook
---

The goal of this file is to put together all of the information that I have collected during the microscopy trials of the HT22 cells. For these trials, the protocol was changed and dropped to 4 days.  There were a few reasons for this. First, the cell growth behavior changed from before spring break to after spring break.  This is because the cells are getting higher in their n numbers (the number of splits).  Typically, the more you split cells, the more their behavior begins to change.  As such, separate growth rate data was collected during these 5 day trials to observe the differences in growth between the control and experimental conditions.

```{r}

# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)
library(readr)

```

Now import the data
```{r}

counts <- read_csv(here("Data", "CMF6_HT22Microscopy_3-27_4-3.csv"))

# take a look at the structure of the data
glimpse(counts)

# recode the counts to be before and after exposure
counts$Condition <- recode(counts$Condition, "B-Field" = 'Magnetic Field')

```
```{r}
CMF7 <- read.csv(here("Data", "CMF7_HT22Microscopy_4-3_4-8.csv"))

# Add in the cmf7 data
counts <- bind_rows(counts, CMF7)
```

Find the summary info from this data:
```{r}
# find the means of the cell counts and put them in a new table
mean_counts <- counts %>% 
  filter(!is.na(Date), !is.na(Trial_Day), !is.na(Cell_Count), Condition == "Control" | Condition == "Magnetic Field") %>%
    group_by(Condition,Trial_Day) %>% summarise(
      meancounts = mean(Cell_Count),
      secounts = sd(Cell_Count)/sqrt(sum(!is.na(Cell_Count)))
       )
```

```{r}
# create scatterplot with error bars
ggplot(mean_counts, aes(Condition, meancounts, color = Trial_Day))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    facet_wrap(~Trial_Day)+
    theme(legend.position="none")+
    xlab("Condition")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
         
```
Now lets make the relationship a little more obvious between the two

```{r}
library(scales)
# create scatter plot with error bars
ggplot(mean_counts, aes(Trial_Day, meancounts, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4', "dodgerblue"))+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    #theme(legend.position="none")+
    xlab("Trial Day")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
```

Calculate the growth rate data:
```{r}
# calculate the growth rates using a for loop and store the results of the data in   a new dataframe

#define the starting value of x (where we want the loop to start iterating through   our previous data set)
  x <- 1

# make an empty data frame
d <- data.frame()

# Write for loop to go through and create a new data set with the growth rates
  for(i in mean_counts$meancounts) {
    y = mean_counts$meancounts[x + 1]
    
    z = (y-mean_counts$meancounts[x])/24
    
    #print(z)
    
    x = x+1
    
    d = rbind(d, z)
  }

dput(d)
```

```{r}
# add in the hours and slope variables to look at how the growth rate changes over time
mean_counts["Hours"] <-c(0,24,48,72, 0,24,48,72)
mean_counts["Slope"] <- c(NA, 1.27170138888889, 3.37239583333333, 
4.5390625, NA, 1.68576388888889, 3.99479166666667, 
5.53645833333333)

```

Great, now that its added into the mean_counts data frame, we can use that to create a graph of the slopes:
```{r}
# create scatter plot that shows the slopes of each of the conditions
ggplot(mean_counts, aes(Hours, Slope, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4'))+
     #geom_errorbar(aes(ymin = standardcounts - secounts,
                       #ymax = standardcounts + secounts), width = 0.1)+
    geom_area(aes(fill = Condition, alpha = 0.2))+
    
    theme_bw()+
    scale_x_continuous(breaks=seq(0,144,by=24))+
    theme(legend.position="none")+
    xlab("Time (Hours)")+
    ylab(expression("Growthrate (Cells/hour)")) +
  facet_grid(~Condition)
```

There were not enough trials performed during this analysis to determine if there is statistical relavence with regards to the growth and proliferation observed in these trials. 

# Move on to microscopy analysis

Here I am going to try and create a polar histogram with the length and angle data from the preliminary analysis of the HT22 microscope images.

First we need to import the data:
```{r}
control_la <- read.csv(here("Data", "CMF6_ControlB_la.csv"))

control_circ <- read.csv(here("Data", "CMF6_ControlB_circ.csv"))

MF_la <- read.csv(here("Data", "CMF6_MF_la.csv"))
  
MF_circ <- read.csv(here("Data", "CMF6_MF_circ.csv"))
```

Clean up the MF_la data and input the data that is scaled correctly
```{r}
MF_la_scaled <- read_csv(here("Data", "CMF6_MFD_Numbers", "CMF6_MFD_la_scaled.csv"))

MF_la <- 
  MF_la %>%
  filter(Length < 200)

bind_rows(MF_la, MF_la_scaled)
```

Now try the analysis:
```{r}
library(plotrix)
polar.plot(control_la$Length, control_la$Angle, lwd = 3, line.col = 4, show.grid.labels=FALSE)
```



Control data: 

```{r}
library(scales)
base <- ggplot(control_la, aes(x=Angle, y=Length))

p <- base + coord_polar(start = 0, direction = -1)+
  scale_x_continuous(limits = c(-180,180), breaks = c(0, 45, -45, 180, 
  90, -90, 135, -135), labels = c("0", "45", "315", "180", "90", "270", "135", "225"))

p+ geom_segment(aes(y=0, xend=Angle, yend=Length), color = '#F8766D')+
  scale_y_continuous(limits = c(0, 120), breaks = c(0, 30, 60, 90, 120),"Neurite Length (\u00b5m)")+
  theme_bw()+
  annotate("text", x = 0, y = 0,
           label = expression("0"))+
  annotate("text", x = -160, y = 30,
           label = expression("30")) +
  annotate("text", x = -160, y = 60,
           label = expression("60"))+
  annotate("text", x = -160, y = 90,
           label = expression("90"))+
  annotate("text", x = -160, y = 120,
           label = expression("120"))
  
```

Magnetic field data:
```{r}
library(scales)
base <- ggplot(MF_la, aes(x=Angle, y=Length))

p <- base + coord_polar(start = 0, direction = -1)+
  scale_x_continuous(limits = c(-180,180), breaks = c(0, 45, -45, 180, 
  90, -90, 135, -135), labels = c("0", "45", "315", "180", "90", "270", "135", "225"))

p+ geom_segment(aes(y=0, xend=Angle, yend=Length), color = "#00BFC4")+
  scale_y_continuous(limits = c(0, 120), breaks = c(0, 30, 60, 90, 120),"Neurite Length(\u00b5m)")+
  theme_bw()+
  annotate("text", x = 0, y = 0,
           label = expression("0"))+
  annotate("text", x = -160, y = 30,
           label = expression("30")) +
  annotate("text", x = -160, y = 60,
           label = expression("60"))+
  annotate("text", x = -160, y = 90,
           label = expression("90"))+
  annotate("text", x = -160, y = 120,
           label = expression("120"))
  
```




```{r}
# calculate the circularity in the circularity data sets
control_circ <-
  control_circ %>%
  mutate(circ = 4*pi *(Area/(Perim.*Perim.)))

MF_circ <-
  MF_circ %>%
  mutate(circ = 4*pi *(Area/(Perim.*Perim.)))

```

```{r}
# combine all of the data
full_circ <- bind_rows("Control" = control_circ,
          "Magnetic Field" = MF_circ,
          .id = "Condition")
```

Now let's graph this:
```{r}
ggplot(full_circ, aes(x = circ, fill = Condition))+
  geom_histogram(bins = 50, alpha = 0.5, position="identity", aes(y=..density.., fill=Condition, ))+
 geom_density(alpha=.2) +
  theme_bw() +
  labs(x = "Circularity bins", y = "Neuron Density")+
  scale_x_continuous(limits = c(0,1), breaks = c(0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 1))

  
  
```
```{r}
circ_plt <- ggplot(full_circ, aes(x = circ, fill = Condition))+
  #geom_histogram(bins = 50, alpha = 0.5, position="identity", aes(y=..density.., fill=Condition, ))+
 geom_density(alpha=.2) +
  theme_bw() +
  labs(x = "Circularity bins", y = "Neuron Density")+
  scale_x_continuous(limits = c(0,1), breaks = c(0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 1))

ggsave(filename = "circ_plt.jpg",
       plot = circ_plt,
       width = 7,
       height = 5)
```

Lets see if we can model the differences in circularity a little differently:
```{r}
mean_circ <- full_circ %>%
  group_by(Condition)%>%
  summarise(meancirc = mean(circ),
            secirc = sd(circ)/sum(!is.na(circ)))
```



Lets model the length and angle data using a density plot as well:
```{r}
# rename the first columns in the control and MF data sets so they will join together
control_la <- control_la %>%
  rename("Line" = 1)

MF_la <- MF_la %>%
  rename("Line" = 1)

# join the mf_la an control_la data frames together to compare them
full_la <- bind_rows("Control" = control_la,
          "Magnetic Field" = MF_la,
          .id = "Condition")

```


Length histograms/density plots:
```{r}
ggplot(full_la, aes(x = Length, fill = Condition))+
  geom_histogram(bins = 50, alpha = 0.5, position="identity", aes(y=..density.., fill=Condition, ))+
 geom_density(alpha=.2) +
  theme_bw() +
  labs(x = "Neurite Length (\u00b5m)", y = "Neuron Density")
  
```

```{r}
length_plt <- ggplot(full_la, aes(x = Length, fill = Condition))+
  #geom_histogram(bins = 50, alpha = 0.5, position="identity", aes(y=..density.., fill=Condition, ))+
 geom_density(alpha=.2) +
  theme_bw() +
  labs(x = "Neurite Length (\u00b5m)", y = "Neuron Density")+
  theme(legend.position = "none")

ggsave(filename = "length_density_plt.jpg",
       plot = length_plt,
       width = 5,
       height = 5)
  
```

Now let's see if there is a significant difference between the lengths of the neurites:
```{r}
# run a t.test to test the difference between the mean neurite length between the two conditions:
lengthtest <- t.test(Length~Condition, data = full_la)
lengthtest
```
Run a two sample t test on the circularity:
```{r}
circtest <- t.test(circ~Condition, data = full_circ)
circtest
```

```{r}
angle_boxplt <- full_la %>%
  ggplot(aes(x= Condition, y =Angle, fill = Condition))+
  geom_boxplot(alpha = 0.5)+
  theme_bw()+
  theme(legend.position="none")
  
ggsave(filename = "angle_boxplt.jpg",
       plot = angle_boxplt,
       width = 5,
       height = 5)
  
```
```{r}
lattest <- t.test(Angle~Condition, data = full_la)
lattest
```


```{r}
full_la %>%
  ggplot(aes(x= Length, fill = Condition))+
  geom_boxplot()


full_la%>%
  group_by(Condition)%>%
  summarise(meanangle = mean(Angle),
            meanlength = length(Length))
```









