---
title: "HT22 Growth Analysis"
output: html_notebook
---

This code is designed to analyze the cell count data from all 3 of the complete experiments that have been done this semester so far.  The goal is to try and use the data from all of the trials to identify potential trends that could provide insight into the growth behavior of the cells. 

General Notes:
- cells seeded at 10,000 cells/plate
- 2 plates running at once (cells are swapped halfway through exposure)

Experimental Protocol:
- Day 0: seeding and growth
- Day 1: counting, exposure
- Day 2: counting, exposure, 1/2 media change
- Day 3: counting, exposure
- Day 4: counting, exposure, 1/2 media change
- Day 5: counting, exposure
- Day 6: counting, exposure, 1/2 media change
- Day 7: counting, disposal

```{r}
# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)
library(readr)
```

Now I will import the data from all of the trials that I have successfully completed:
```{r}
counts <- read_csv(here("Data", "CMF_HT22_Cummulative_analysis.csv"))

# take a look at the structure of the data
glimpse(counts)

# recode the counts to be before and after exposure
counts$Condition <- recode(counts$Condition, "B-Field" = 'Magnetic Field')

# relevel the data to put the control first
#counts<- mutate(counts, Condition = relevel(Condition, ref = "Control"))
```

```{r}
# find the means of the cell counts and put them in a new table
mean_counts <- counts %>% 
  filter(!is.na(Date), !is.na(Trial_Day), Condition == "Control" | Condition == "Magnetic Field") %>%
    group_by(Condition,Trial_Day) %>% summarise(
      meancounts = mean(Cell_Count),
      secounts = sd(Cell_Count)/sqrt(sum(!is.na(Cell_Count)))
  
       )
```
Great, now lets get to work with this data

```{r}
# create scatterplot with error bars
ggplot(mean_counts, aes(Condition, meancounts, color = Trial_Day))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    facet_wrap(~Trial_Day)+
    theme(legend.position="none")+
    xlab("Condition")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
         
```

Now lets make the relationship a little more obvious between the two

```{r}
library(scales)
# create scatter plot with error bars
ggplot(mean_counts, aes(Trial_Day, meancounts, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4', "dodgerblue"))+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    #theme(legend.position="none")+
    xlab("Trial Day")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
```

