---
title: "CMF HT22 Microscopy Analysis"
output: html_notebook
---

The goal of this file is to put together all of the information that I have collected during the microscopy trials of the HT22 cells. For these trials, the protocol was changed and dropped to 4 days.  There were a few reasons for this. First, the cell growth behavior changed from before spring break to after spring break.  This is because the cells are getting higher in their n numbers (the number of splits).  Typically, the more you split cells, the more their behavior begins to change.  As such, seperate growth rate data was collected during these 5 day trials to observe the differences in growth between the control and experimental conditions.

```{r}

# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)
library(readr)

```

Now import the data
```{r}

counts <- read_csv(here("Data", "CMF6_HT22Microscopy_3-27_4-3.csv"))

# take a look at the structure of the data
glimpse(counts)

# recode the counts to be before and after exposure
counts$Condition <- recode(counts$Condition, "B-Field" = 'Magnetic Field')

```
```{r}
CMF7 <- read.csv(here("Data", "CMF7_HT22Microscopy_4-3_4-8.csv"))

# Add in the cmf7 data
counts <- bind_rows(counts, CMF7)
```

Find the summary info from this data:
```{r}
# find the means of the cell counts and put them in a new table
mean_counts <- counts %>% 
  filter(!is.na(Date), !is.na(Trial_Day), !is.na(Cell_Count), Condition == "Control" | Condition == "Magnetic Field") %>%
    group_by(Condition,Trial_Day) %>% summarise(
      meancounts = mean(Cell_Count),
      secounts = sd(Cell_Count)/sqrt(sum(!is.na(Cell_Count)))
       )
```

```{r}
# create scatterplot with error bars
ggplot(mean_counts, aes(Condition, meancounts, color = Trial_Day))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    facet_wrap(~Trial_Day)+
    theme(legend.position="none")+
    xlab("Condition")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
         
```
Now lets make the relationship a little more obvious between the two

```{r}
library(scales)
# create scatter plot with error bars
ggplot(mean_counts, aes(Trial_Day, meancounts, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4', "dodgerblue"))+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    #theme(legend.position="none")+
    xlab("Trial Day")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
```

Calculate the growth rate data:
```{r}
# calculate the growth rates using a for loop and store the results of the data in   a new dataframe

#define the starting value of x (where we want the loop to start iterating through   our previous data set)
  x <- 1

# make an empty data frame
d <- data.frame()

# Write for loop to go through and create a new data set with the growth rates
  for(i in mean_counts$meancounts) {
    y = mean_counts$meancounts[x + 1]
    
    z = (y-mean_counts$meancounts[x])/24
    
    #print(z)
    
    x = x+1
    
    d = rbind(d, z)
  }

dput(d)
```

```{r}
# add in the hours and slope variables to look at how the growth rate changes over time
mean_counts["Hours"] <-c(0,24,48,72, 0,24,48,72)
mean_counts["Slope"] <- c(NA, 1.27170138888889, 3.37239583333333, 
4.5390625, NA, 1.68576388888889, 3.99479166666667, 
5.53645833333333)

```

Great, now that its added into the mean_counts data frame, we can use that to create a graph of the slopes:
```{r}
# create scatter plot that shows the slopes of each of the conditions
ggplot(mean_counts, aes(Hours, Slope, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4'))+
     #geom_errorbar(aes(ymin = standardcounts - secounts,
                       #ymax = standardcounts + secounts), width = 0.1)+
    geom_area(aes(fill = Condition, alpha = 0.2))+
    
    theme_bw()+
    scale_x_continuous(breaks=seq(0,144,by=24))+
    theme(legend.position="none")+
    xlab("Time (Hours)")+
    ylab(expression("Growthrate (Cells/hour)")) +
  facet_grid(~Condition)
```


# Move on to microscopy analysis

Here I am going to try and create a polar histogram with the length and angle data from the preliminary analysis of the HT22 microscope images.

First we need to import some sample data:
```{r}
circ <- read.csv(here("Data", "Control1_la_results.csv"))
```

Now try the analysis:
```{r}
ggplot(circ, aes(x = Length)) +
  geom_histogram(binwidth = 10) +
  scale_x_continuous(breaks = seq(0, 360, 60)) +
  coord_polar() +
  xlab(NULL)+ylab(NULL)

library(plotrix)
polar.plot(circ$Length, circ$Angle, lwd = 3, line.col = 4, show.grid.labels=FALSE)
```

ggplot(iris, aes(x = Sepal.Width)) +
  geom_histogram(binwidth = .1) +
  scale_x_continuous(breaks = seq(0, 360, 60)) +
  coord_polar() +
  xlab(NULL)+ylab(NULL)
  
New trial with plotting using `plotly`
```{r}
library(plotly)


```

```{r}
circ <-
  circ %>%
  mutate(Angle_Rad = Angle/57.2957795)
```

```{r}
ggplot(circ, aes(x = Angle_Rad)) + 
  coord_polar(theta = "x", start = 2*pi, direction = -1)+
  geom_histogram(aes(fill = Length), color= "black") +
  scale_y_continuous("Length of Neurite (microns)", limits = c(0,100)) +
  scale_x_continuous(limits = c(-pi,pi), breaks = c(0, pi/4, pi/2, 3*pi/4, 
  pi, -3*pi/4, -pi/2, -pi/4),
                     labels = c("270", "315", "0", "45", "90", "135", "180", "225"))
```
```{r}
base <- ggplot(circ, aes(x=Angle, y=Length))

p <- base + coord_polar(start = 0, direction = -1)+
  scale_x_continuous(limits = c(-180,180), breaks = c(0, 45, -45, 180, 
  90, -90, 135, -135), labels = c("0", "45", "90", "135", "180", "225", "270", "315"))

p+ geom_segment(aes(y=0, xend=Angle, yend=Length))+
  scale_y_continuous("Length of Neurite (microns)")+
  theme_bw()
  
```



