---
title: "HT22 Growth Analysis"
output: html_notebook
---

This code is designed to analyze the cell count data from all 3 of the complete experiments that have been done this semester so far.  The goal is to try and use the data from all of the trials to identify potential trends that could provide insight into the growth behavior of the cells. 

General Notes:
- cells seeded at 10,000 cells/plate
- 2 plates running at once (cells are swapped halfway through exposure)

Experimental Protocol:
- Day 0: seeding and growth
- Day 1: counting, exposure
- Day 2: counting, exposure, 1/2 media change
- Day 3: counting, exposure
- Day 4: counting, exposure, 1/2 media change
- Day 5: counting, exposure
- Day 6: counting, exposure, 1/2 media change
- Day 7: counting, disposal

```{r}
# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)
library(readr)
```

Now I will import the data from all of the trials that I have successfully completed:
```{r}
counts <- read_csv(here("Data", "CMF_HT22_Cummulative_analysis.csv"))

# take a look at the structure of the data
glimpse(counts)

# recode the counts to be before and after exposure
counts$Condition <- recode(counts$Condition, "B-Field" = 'Magnetic Field')

```


```{r}
# find the means of the cell counts and put them in a new table
mean_counts <- counts %>% 
  filter(!is.na(Date), !is.na(Trial_Day), !is.na(Cell_Count), Condition == "Control" | Condition == "Magnetic Field") %>%
    group_by(Condition,Trial_Day) %>% summarise(
      meancounts = mean(Cell_Count),
      secounts = sd(Cell_Count)/sqrt(sum(!is.na(Cell_Count)))
  
       )
```
Great, now lets get to work with this data

```{r}
# create scatterplot with error bars
ggplot(mean_counts, aes(Condition, meancounts, color = Trial_Day))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    facet_wrap(~Trial_Day)+
    theme(legend.position="none")+
    xlab("Condition")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
         
```

Now lets make the relationship a little more obvious between the two

```{r}
library(scales)
# create scatter plot with error bars
ggplot(mean_counts, aes(Trial_Day, meancounts, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4', "dodgerblue"))+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    #theme(legend.position="none")+
    xlab("Trial Day")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
```
Now let's put together the growth rate data for each of these trials:
```{r}
# calculate the growth rates using a for loop and store the results of the data in   a new dataframe

#define the starting value of x (where we want the loop to start iterating through   our previous data set)
  x <- 1

# make an empty data frame
d <- data.frame()

# Write for loop to go through and create a new data set with the growth rates
  for(i in mean_counts$meancounts) {
    y = mean_counts$meancounts[x + 1]
    
    z = (y-mean_counts$meancounts[x])/24
    
    #print(z)
    
    x = x+1
    
    d = rbind(d, z)
  }

dput(d)
```

```{r}
# add in the hours and slope variables to look at how the growth rate changes over time
mean_counts["Hours"] <-c(0,24,48,72,96, 120, 144, 0,24,48,72,96, 120, 144)
mean_counts["Slope"] <- c(NA, 0.015625, 0.396412037037037, 1.91840277777778, 
1.91782407407407, 3.48611111111111, 3.88946759259259, NA, 
0.774028784219002, 0.613309178743962, 2.58256944444444, 3.33275462962963, 
4.24016203703704, 1.29398148148148)

```

Great, now that its added into the mean_counts data frame, we can use that to create a graph of the slopes:
```{r}
# create scatter plot that shows the slopes of each of the conditions
ggplot(mean_counts, aes(Hours, Slope, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4'))+
     #geom_errorbar(aes(ymin = standardcounts - secounts,
                       #ymax = standardcounts + secounts), width = 0.1)+
    geom_area(aes(fill = Condition, alpha = 0.2))+
    
    theme_bw()+
    scale_x_continuous(breaks=seq(0,144,by=24))+
    theme(legend.position="none")+
    xlab("Time (Hours)")+
    ylab(expression("Growthrate (Cells/hour)")) +
  facet_grid(~Condition)
```

```{r}
# create scatter plot that shows the slopes of each of the conditions
ggplot(mean_counts, aes(Hours, Slope, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4'))+

    #geom_area(aes(fill = Condition, group = Condition), alpha = 0.2)+
    
    theme_bw()+
    scale_x_continuous(breaks=seq(0,144,by=24))+
    theme(legend.position="none")+
    xlab("Time (Hours)")+
    ylab(expression("Growthrate (Cells/hour)")) 
  
```