---
title: "HT22 Cell count analysis 2-10-23 to 2-13-23 "
output: html_notebook
---
This code is designed to analyze the cell count data from the first experimental trial after putting mouse hippocampal cells in the EMF for 30 mins.  The goal of this trial is to measure survivalbility of these cells and ensure that the current protocol won't kill these neural cells before we try with C17.2 stem cells.

These cells were seeded at 25,000 cells/plate and this was the first trial of 2023 after working out a few kinks with the equipment

```{r}
# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)

```

Import the data:
```{r}
# import the data set that we will be analysing, the stringsAsFactors = True argument converts categorical data to factors so we can analyse the levels
counts <- read.csv(here("Data", "HT22cellcountdata_2-11_2-13.csv"),stringsAsFactors = TRUE)

# take a look at the structure of the data
glimpse(counts)

# recode the counts to be before and after exposure
counts$Condition <- recode(counts$Condition, "B-Field" = 'Magnetic Field')

# relevel the data to put the control first
counts<- mutate(counts, Condition = relevel(Condition, ref = "Control"))
```

Now I am going to make a boxplot of the data to visualize the distributions
better idea of the distribution.

```{r}
# make a box plot of the data
ggplot(counts, aes(x = Condition, y = Cell_Count)) +
geom_boxplot()+
  geom_point(size = 3, colour = 'lightgrey', alpha = 0.5)+
  facet_wrap(~Trial_Day)+
  xlab("Sample") +
  ylab("Cell counts (cells/mm^2)")
```

The boxplot doesn't look the best but right now it is taking all of the count data and showing the distribution.  A more accurate teller of the actual data will be if we take the average of the counts by day.

```{r}
# find the means of the cell counts and put them in a new table
mean_counts <- counts %>% group_by(Condition,Trial_Day) %>% summarise(
  meancounts = mean(Cell_Count),
  secounts = sd(Cell_Count)/sqrt(sum(!is.na(Cell_Count))                                                             ))
```

Great, now lets get to work with this data

```{r}
# create scatterplot with error bars
ggplot(mean_counts, aes(Condition, meancounts, color = Trial_Day))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    facet_wrap(~Trial_Day)+
    theme(legend.position="none")+
    xlab("Condition")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
         
```

Now lets make the relationship a little more obvious between the two

```{r}
library(scales)
# create scatter plot with error bars
high_density_plt <- ggplot(mean_counts, aes(Trial_Day, meancounts, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     scale_color_manual(values=c('#F8766D','#00BFC4'))+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    #theme(legend.position="none")+
    xlab("Trial Day")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))

ggsave(filename = "high_density_plt_sem2.jpg",
       plot = high_density_plt,
       width = 7,
       height = 4)
```
