---
title: "HT22 Mouse Hippocampal Cell counts 11-16"
output: html_notebook
---

This analysis is for the HT22 mouse hippocampal trials from 11-16 to 11-19.  During this trial, the cells were seeded at 25,000 cells per plate, grown for one day, and exposed to the magnetic field as normal.  The evening of the day of magnetic field exposure, the cell media was changed using a normal glucose DMEM.  

We found out later that HT22 cells do not really like normal glucuse DMEM as it accelerated glutamate induced toxicity in the cells.  Nevertheless I will proceed with the analysis because if we can show that the B-Field cells died to the same degree or less than the control condition it still supports our hypothesis.

First import the data:
```{r}
# clear the environment
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)

cells <- read.csv(here("Data", "Cellcountdata_11-9_11-13.csv"),stringsAsFactors = TRUE)
```

Now let's make a quick boxplot to take a look at the data:

```{r}
# relevel the data to put the control first
cells<- mutate(cells, Condition = relevel(Condition, ref = "Control"))

# reassign the values of the different EMF conditions to combine them
cells <- cells %>% mutate(Exposure = recode(cells$Date, "11/10/2022" = '24 Hours After Exposure', "11/9/2022" = 'Before Exposure'))

# relevel the data to put the before exposure first
cells<- mutate(cells, Exposure = relevel(Exposure, ref = "Before Exposure"))

cells <- cells %>% mutate(Condition = recode(cells$Condition, "Control" = 'Control', "B-Field" = 'Magnetic Field'))


# make a box plot of the data
ggplot(cells, aes(x = Condition, y = Cells)) +
geom_boxplot()+
  geom_point(size = 3, colour = 'lightgrey', alpha = 0.5)+
  facet_wrap(~Exposure)+
  xlab("Sample") +
  ylab("Cell counts (cells/mm^2)")
```

This is looking good except, I'd like to put before on the left and after on the right...

```{r}
# find the means of the cell counts and put them in a new table
mean_counts <- cells %>% group_by(Condition,Exposure) %>% summarise(
  meancounts = mean(Cells),
  secounts = sd(Cells)/sqrt(sum(!is.na(Cells))                                                             ))
```
Now lets make a scatterplot of this:
```{r}
# create scatterplot with error bars
ggplot(mean_counts, aes(Condition, meancounts, color = Exposure))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    facet_wrap(~Exposure)+
    theme(legend.position="none")+
    xlab("Condition")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))
```


```{r}
# create scatter plot with error bars
media_change_plot <- ggplot(mean_counts, aes(Exposure, meancounts, color = Condition, group = Condition))+
     geom_point()+
     geom_line()+
     geom_errorbar(aes(ymin = meancounts - secounts,
                       ymax = meancounts + secounts), width = 0.1)+
    theme_bw()+
    #theme(legend.position="none")+
    xlab("Count")+
    ylab(expression(Cell~Counts~(cells~"/"~mm^2)))

media_change_plot

ggsave(filename = "media_change_plt_sem1.jpg",
       plot = media_change_plot,
       width = 7,
       height = 5)

```

