---
title: "Helmholtz Coil MF Exposure System Control Testing"
author: "Cora Ferguson"
date: "2023-01-24"
output: html_document
---

# Overview
Today the goal is to determine if the system needs to be recalibrated from when I last used it last semester.  Testing consisted of me using my phone and the app *Phyphox* to measure the magnitude of the magnetic field in four different conditions:

Magnetic Field | Arduino System
--------------- | ---------------
On | On
On | Off
Off |On
Off | Off

The initial data makes it look like there aren't any significant shifts from last semester but I'll plot these to take a closer look. 

# Preliminary Data Analysis
```{r}
# set up the environment and import the data
rm(list = ls())

# import the libraries that we need
library(dplyr)
library(here)
library(ggplot2)
library(readr)
```

```{r}
# import the data
DF <- read.csv(here("Data","SystemControlTesting_1-24-2023.csv"), stringsAsFactors = TRUE)

glimpse(DF)
```

```{r}
# make a new data set taking the averages of the different conditions
avg_DF <- DF %>% group_by(Distance, Test) %>% summarise(
     meanMagnitude = mean(Magnitude_uT),
     seMagnitude = sd(Magnitude_uT)/sqrt(sum(!is.na(Magnitude_uT))))
```

Now that things are a little cleaned up, I will graph them using scatter plots:
```{r}
# filter things down to test A (Magnetic field off, Arduino off)
 avg_DFA<- filter(avg_DF, Test == "A")

ggplot(avg_DFA, aes(Distance, meanMagnitude))+
         geom_point()+
     xlab("Distance from center of coils (cm)")+
     ylab("Magnetic field (uT)")+
     theme_bw()
  

ggplot(avg_DFA, aes(Distance, meanMagnitude))+
         geom_point()+
     xlab("Distance from center of coils (cm)")+
     ylab("Magnetic field (uT)")+
     theme_bw()+
  ylim(0, 1000)
```

```{r}
# filter things down to test B (Magnetic field off, Arduino on)
 avg_DFB<- filter(avg_DF, Test == "B")

ggplot(avg_DFB, aes(Distance, meanMagnitude))+
         geom_point()+
     xlab("Distance from center of coils (cm)")+
     ylab("Magnetic field (uT)")+
     theme_bw()
  

ggplot(avg_DFB, aes(Distance, meanMagnitude))+
         geom_point()+
     xlab("Distance from center of coils (cm)")+
     ylab("Magnetic field (uT)")+
     theme_bw()+
  ylim(0, 1000)
```


```{r}
# filter things down to test C (Magnetic field on, Arduino off)
 avg_DFC<- filter(avg_DF, Test == "C")

ggplot(avg_DFC, aes(Distance, meanMagnitude))+
         geom_point()+
     xlab("Distance from center of coils (cm)")+
     ylab("Magnetic field (uT)")+
     theme_bw()
  

ggplot(avg_DFC, aes(Distance, meanMagnitude))+
         geom_point()+
     xlab("Distance from center of coils (cm)")+
     ylab("Magnetic field (uT)")+
     theme_bw()+
  ylim(0, 1200)
```

```{r}
# filter things down to test C (Magnetic field on, Arduino on)
 avg_DFD<- filter(avg_DF, Test == "D")

ggplot(avg_DFD, aes(Distance, meanMagnitude))+
         geom_point()+
     xlab("Distance from center of coils (cm)")+
     ylab("Magnetic field (uT)")+
     theme_bw()
  

ggplot(avg_DFD, aes(Distance, meanMagnitude))+
         geom_point()+
     xlab("Distance from center of coils (cm)")+
     ylab("Magnetic field (uT)")+
     theme_bw()+
  ylim(0, 1200)
```
